### 2020-05-12 response.tex

\newcommand{\comment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
\newcommand{\jd}[1]{\comment{cyan}{JD}{#1}}
\newcommand{\swp}[1]{\comment{magenta}{SWP}{#1}}
\newcommand{\dc}[1]{\comment{blue}{DC}{#1}}
\newcommand{\jsw}[1]{\comment{green}{JSW}{#1}}
\newcommand{\hotcomment}[1]{\comment{red}{HOT}{#1}}

\begin{document}

\noindent Dear Editor:

Thank you for the chance to revise and resubmit our manuscript. 
We have re-written a considerable amount of the manuscript in order to convince a more general readership (suggested by reviewer 1) and to provide a clearer explanation of the methods (suggested by reviewer 2).
A large proportion of the Results section has been moved to the Discussion section, as suggested by reviewer 2.
We have also included a new paragraph in the Discussion section to compare early estimates  of the basic reproductive number with more recent estimates.
We had previously interpreted the uncertainty range reported by Study 2 as a 95\% credible interval; we now interpret it as 90\% credible interval as it better suits their definition.
Otherwise, the main figures and the overall conclusions remain unchanged.
Below please find our responses to reviewers.

### 2020-06-21 backup/ncov_b1.tex
done

\usepackage[numbers,sort&compress]{natbib}
\bibliographystyle{unsrtnat}
\date{\today}

@@

\section*{Abstract}

\subsection*{Background}

A novel coronavirus (2019-nCoV) has recently emerged as a global threat.
As the epidemic progresses, many disease modelers have focused on estimating the basic reproductive number \Ro, the average number of secondary cases caused by a primary case in an otherwise susceptible population.
The modeling approaches and resulting estimates of \Ro vary widely, despite relying on similar data sources.

\subsection*{Aim}

We aimed to develop a framework for comparing and combining different estimates of \Ro across a wide range of models.

\subsection*{Methods}

We reviewed 7 model-based analyses of the 2019-nCoV outbreak that were published online between January 23--26, 2020.
We decompose their \Ro estimates into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
We use a Bayesian multilevel model to construct pooled estimates and measure uncertainties associated with these quantities.

\subsection*{Results}

We find that most early estimates of \Ro rely on strong assumptions, especially about the generation-interval dispersion.
Estimates that rely on narrow generation-interval distributions are overly sensitive to estimates of the exponential growth rate.

\subsection*{Conclusion}

Our results emphasize the importance of propagating uncertainties in all components of \Ro, including the shape of the generation-interval distribution, in efforts to estimate \Ro at the outset of an epidemic.

@@

Basic reproductive number, 2019-nCoV, novel coronavirus, Bayesian multilevel model

\section*{Declaration of interests}

@@

Since December 2019 \cite{pneumonia}, a novel coronavirus (2019-nCoV) has been spreading in China and other parts of the world.
Although the virus is believed to have originated from animal reservoirs, the ability of 2019-nCoV to directly transmit between humans has posed a greater threat for its spread \cite{chan2020familial}.
As of February 3, 2020, the World Health Organization (WHO) has confirmed 17391 cases, including 153 confirmed cases in 23 different countries, outside China \cite{who02report}.
% Also on January 30, 2020, WHO declared the outbreak a Public Health Emergency of International Concern \cite{whoemer}.

As the disease continues to spread, many researchers have published their preliminary analyses of the outbreak as pre-prints \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov, riouncov}, focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \cite{anderson1991infectious}).
Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission, and about the final size of the outbreak \cite{anderson1991infectious}.
We commend these researchers for their timely contribution and those who made the data publicly available.
However, it can be difficult to compare a disparate set of estimates of \Ro from different research groups when the estimation methods and their underlying assumptions vary widely.

Here, we show that a wide range of approaches to estimating \Ro can be understood and compared in terms of estimates of three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
The generation interval, defined as the interval between the time when an individual becomes infected and the time when that individual infects another individual \cite{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro \cite{wallinga2007generation};
therefore, estimates of \Ro from different models directly depend on their implicit assumptions about the generation-interval distribution and the exponential growth rate.
Early in an epidemic, information is scarce and there is inevitably a great deal of uncertainty surrounding disease transmission.
We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty in \Ro should propagate error from all three sources \cite{elderd2006uncertainty}.

We compare seven disparate models published online between January 23--26, 2020 that estimated \Ro for the 2019-nCoV outbreak \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}. 
We use our framework to construct pooled estimates for the three key quantities: $r$, $\bar G$, and $\kappa$;
the pooled estimates reflect the uncertainties present in modeling approaches.
We use these pooled estimates to illustrate the importance of propagating different sources of error, particularly uncertainty in both the growth rate and the generation interval.
We also use our framework to tease apart which assumptions of these different models led to their different estimates and confidence intervals. 
Despite the availability of more recent and/or updated estimates of \Ro, we restrict ourselves to the estimates above in order to focus on the resolution of uncertainty in the earliest stages of an epidemic.

@@

\begin{table}[t]
\begin{center}
\scriptsize
\begin{tabular}{l|p{3.5cm}|p{2.5cm}|p{2.7cm}|l}
 & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ & \\
\hline
Study 1 & 1.5--3.5 & 10 & 1 & Bedford \textit{et al.} \cite{bedfordncov} \\
\hline
Study 2 & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ & Imai \textit{et al.} \cite{imaincov} \\
\hline
Study 3 & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 & Liu \textit{et al.} \cite{liuncov} \\
\hline
Study 4 & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & Read \textit{et al.} \cite{readncov} \\
\hline
Study 5 & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5 & Riou and Althaus \cite{riouncov} \\
\hline
Study 6 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2 & Zhao \textit{et al.} \cite{zhaoncov} \\
\hline
Study 7 & 2.0--3.1 & 6--10 & 0 & Majumder and Mandl \cite{majumderncov} \\
\hline
\end{tabular}
\end{center}
\caption{
\textbf{Reported estimates of the basic reproductive number and the assumptions about the generation-interval distributions.}
Estimates of \Ro and their assumptions about the shape of the generation interval distributions were collected from 7 studies.
$^\ast$We treat these intervals as a 95\% confidence interval in our analysis.
$^\dagger$We assume $\kappa = 0.5$ in our analysis.
$^\ddagger$The authors presented \Ro estimates under different assumptions regarding the reporting rate; we use their baseline scenario in our analysis to remain consistent with other studies, which do not account for changes in the reporting rate.
}
\end{table}

We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 7 articles that were published online between January 23--26, 2020 (Table 1).
Five studies \cite{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the web site of Imperial College London \cite{imaincov}; and one report was posted on nextstrain.org \cite{bedfordncov}.
Their modeling approaches vary widely: a branching process model \cite{bedfordncov, imaincov, riouncov}, a deterministic Susceptible-Exposed-Infected-Recovered (SEIR) model \cite{readncov}, an exponential growth model \cite{zhaoncov}, a Poisson offspring distribution model \cite{liuncov}, and the Incidence Decay and Exponential Adjustment (IDEA) model \cite{majumderncov}.
Four studies estimated \Ro by directly fitting their models to incidence data \cite{readncov, zhaoncov, liuncov, majumderncov}.
The remaining three studies estimated \Ro by comparing the predicted number of cases from their models with the estimated number of total cases by January 18 (between 1,000 and 9,7000 \cite{imaincov0})
Some of these studies have now been published in peer-reviewed journals \cite{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \cite{readncov2}.

@@

Early in an outbreak, \Ro is difficult to estimate directly;
instead, \Ro is often inferred from the exponential growth rate $r$, which can be estimated reliably from incidence data \cite{ma2014estimating}.
Given an estimate of the exponential growth rate $r$ and an \emph{intrinsic} generation-interval distribution $g(\tau)$ \cite{champredon2015intrinsic}, the basic reproductive
number can be estimated via the Euler-Lotka equation \cite{wallinga2007generation}:
\begin{linenomath*}
\begin{equation}
1/\Ro = \int \exp(-r\tau) g(\tau) \, d\tau.
\label{eq:euler}
\end{equation}
\end{linenomath*}
In other words, estimates of \Ro must
depend on the assumptions about the
exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.

Here, we use the gamma approximation framework \cite{park2019practical} to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
Assuming that generation intervals follow a gamma distribution
with the mean $\bar G$ and the squared coefficient of variation $\kappa$, 
we have
\begin{linenomath*}
\begin{equation}
\Ro = \left(1 + \kappa r \bar{G}\right)^{1/\kappa}.
\label{eq:gamma}
\end{equation}
\end{linenomath*}
This equation demonstrates that a generation-interval distribution
that has a larger mean (higher $\bar{G}$) or is less variable (lower $\kappa$)
will give a higher estimate of \Ro for the same value of $r$.

@@

\newcommand{\gammdist}{\mathrm{Gamma}}
\begin{table}[t]
\begin{center}
\scriptsize
\begin{tabular}{l|p{4.5cm}|p{2.5cm}|p{2.7cm}}
 & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ \\
\hline
Study 1 & $\mathrm{Uniform}(1.5, 3.5)$ & 10 & 1 \\
\hline
Study 2 & $\gammdist(\mathrm{mean}=2.6, \alpha=28)$ & 8.4 & 0.5 \\
\hline
Study 3 & $\gammdist(\mathrm{mean}=2.92, \alpha=67)$ & 8.4 & 0.2 \\
\hline
Study 4 & $\gammdist(\mathrm{mean}=3.8, \alpha=1400)$ & 7.6 & 0.5 \\
\hline
Study 5 & $\gammdist(\mathrm{mean}=2.2, \alpha=12)$ & $\mathrm{Uniform}(7, 14)$ & 0.5\\
\hline
Study 6 & $\gammdist(\mathrm{mean}=5.47, \alpha=54)$ & $\mathrm{Uniform}(7.6, 8.4)^\ast$ & 0.2\\
\hline
Study 7 & $\exp(r \bar G)^\dagger$ & $\mathrm{Uniform}(6, 10)$ & 0\\
\hline
\end{tabular}
\end{center}
\caption{
\textbf{Probability distributions for \Ro, $\bar G$, and $\kappa$.}
We use these probability distributions to obtain a probability distribution for the exponential growth rate $r$.
The gamma distribution is parameterized by its mean and shape $\alpha$.
Constant values are fixed according to Table 1.
$^\ast$We do not account for this uncertainty during our recalculation of the exponential growth rate $r$ because the reported estimate of $\mathcal R_0$ and its uncertainty assumes $\bar G = 8$; the original article reports three $\mathcal R_0$ (and 95\% CIs) estimates using three different values of $\bar G$: 7.6 (MERS-like), 8 (average), and 8.4 (SARS-like).
We still account for this uncertainty in our pooled estimates ($\mu_G$).
$^\dagger$Study 6 uses the IDEA model \cite{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
Instead of modeling \Ro with a probability distribution and recalculating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in the reported \Ro, when combined with the considered range of $\bar G$.
}
\end{table}

@@

We use weakly informative priors on hyperparameters:
\begin{linenomath*}
\begin{equation}
\begin{aligned}
\mu_r &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week}^{-1},\,\mathrm{shape}=2)\\
\mu_G &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week},\,\mathrm{shape}=2)\\
\mu_\kappa &\sim \gammdist(\mathrm{mean}=0.5,\,\mathrm{shape}=2)\\
(\sigma_r, \sigma_G, \sigma_\kappa) &\sim \textrm{half-normal}(0, 10).
\end{aligned}
\end{equation}
\end{linenomath*}
We followed recommendations outlined in Gelman \textit{et al.} \cite{gelman2006prior}, parameterizing the top-level gamma distributions in terms of their means and standard deviations and imposing weakly informative prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
We had initially used gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) but found this put too much prior probability on large between-study variances. This phenomenon is a known problem \cite{gelman2006prior}.
Alternative choices of prior for the between-study shape parameters are also suboptimal: imposing strong priors (e.g. half-$t(\mu=0,\sigma=1,\nu=4)$  assumes \textit{a priori} that between-study variance is large,  while weak priors (e.g. half-Cauchy(0,5)) can lead to poor mixing.

We run 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 burnin steps and 500,000 sampling steps.
Posterior samples are thinned every 1000 steps.
Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \cite{gelman1992inference};
trace plots and marginal posterior distribution plots are presented in Appendix.
95\% confidence intervals are calculated by taking 2.5\% and 97.5\% quantiles from the marginal posterior distribution for each parameter.

@@

\fref{assumption} compares the reported values of the exponential growth rate $r$, mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ from different studies with the pooled estimates that we calculate from our multilevel model.
We find that there is a large uncertainty associated with the underlying parameters;
many models rely on stronger assumptions that ignore these uncertainties.
Surprisingly, no studies take into account how the variation in generation intervals affects their estimates of \Ro:
all studies assumed fixed values for $\kappa$, ranging from 0 to 1.
% Assuming fixed parameter values can lead to overly strong conclusions \cite{elderd2006uncertainty}.

@@

Estimating the basic reproductive number \Ro is crucial for predicting the course of an outbreak and planning intervention strategies.
Here, we use a gamma approximation \cite{park2019practical} to decompose \Ro estimates into three key quantities ($r$, $\bar G$, and $\kappa$) and apply a multilevel Bayesian framework to compare estimates of \Ro for the novel coronavirus outbreak.
Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, including uncertainties in the amount of dispersion in the generation intervals:
our analysis of individual studies shows that assuming too narrow a generation-interval distribution can make the estimate of \Ro overly sensitive to the estimates of the exponential growth rate $r$.

Of the seven studies that we reviewed, two of them directly fit their models to cumulative number of confirmed cases.
This approach can be appealing because of its simplicity and apparent robustness, but fitting a model to cumulative incidence instead of raw incidence can both bias parameters and give overly narrow confidence intervals, if the resulting non-independent error structures are not taken into account \cite{ma2014estimating, king2015avoidable}.
Naive fits to cumulative incidence data should therefore be avoided.

Many sources of noise affect real-world incidence data, including both dynamical, or ``process'', noise (randomness that directly or indirectly affects disease transmission); and observation noise (randomness underlying how many of the true cases are reported).  
Disease modelers face the choice of incorporating one or both of these in their data-fitting and modeling steps. 
This is not always a serious problem, particularly if the goal is inferring parameters rather than directly making forecasts \cite{ma2014estimating}.
Modelers should however be aware of the possibility that ignoring one kind of error can give overly narrow confidence intervals \cite{king2015avoidable,taylor2016stochasticity}.

There are other important phenomena not covered by our simple framework. 
Examples that seem relevant to this outbreak include: changing reporting rates, reporting delays (including the effects of weekends and holidays), and changing generation intervals.
For emerging pathogens such as 2019-nCoV, there may be an early period of time when the reporting rate is very low due to limited awareness or diagnostic resources;
for example, Zhao \textit{et al.} \cite{zhaoncov} (Study 6) demonstrated that estimates of \Ro can change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 17, when the official diagnostic guidelines were released \cite{who17protocol}, and January 20.
Delays between key epidemiological timings (e.g., infection, symptom onset, and detection) can also shift the shape of an observed epidemic curve and, therefore, affect parameter estimates as well as predictions of the course of an outbreak \cite{tariq2019assessing}.
% Even though a constant delay between infection and detection may not affect the estimate of the growth rate, it can still affect the associated confidence intervals.
Finally, generation intervals can become shorter throughout an epidemic as intervention strategies, such as quarantine, can reduce the infectious period \cite{hethcote2002effects}.
Accounting for these factors is crucial for making accurate inferences.

>>>>

\usepackage{natbib}
\bibliographystyle{chicago}
\date{\today}

@@

\section*{Abstract}
A novel coronavirus (SARS-CoV-2) has recently emerged as a global threat. 
As the epidemic progresses, many disease modelers have focused on estimating the basic reproductive number \Ro -- the average number of secondary cases caused by a primary case in an otherwise susceptible population.
The modeling approaches and resulting estimates of \Ro vary widely, despite relying on similar data sources.
Here, we present a novel statistical framework for comparing and combining different estimates of \Ro across a wide range of models by decomposing the basic reproductive number into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
We then apply our framework to early estimates of \Ro for the SARS-CoV-2 outbreak.
We show that many early \Ro estimates are overly confident.
Our results emphasize the importance of propagating uncertainties in all components of \Ro, including the shape of the generation-interval distribution, in efforts to estimate \Ro at the outset of an epidemic.

@@

SARS-CoV-2, COVID-19, novel coronavirus, basic reproductive number, generation interval, Bayesian multilevel model

@@

Since December 2019, a novel coronavirus (SARS-CoV-2) has been spreading in China and other parts of the world \citep{pneumonia}.
Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, the ability of SARS-CoV-2 ability to directly transmit between humans has posed a greater threat for its spread \citep{huang2020clinical,who26report}.
As of February 27, 2020, the World Health Organization (WHO) has confirmed 82,294 cases of the coronavirus disease (COVID-19), including 3,664 confirmed cases in 46 different countries, outside China \citep{who38report}.
% Also on January 30, 2020, WHO declared the outbreak a Public Health Emergency of International Concern \citep{whoemer}.

As the disease continues to spread, many researchers have already published their analyses of the outbreak as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}) and in peer-reviewed journals (e.g., \cite{li2020early, riou2020pattern, wu2020nowcasting, zhao2020preliminary}), focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the potential final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
We commend these researchers for their timely contribution and those who made the data publicly available.
However, it can be difficult to compare a disparate set of estimates of \Ro from different research groups (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely.

Here, we show that a wide range of approaches to estimating \Ro can be understood and compared in terms of estimates of three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
The generation interval, defined as the interval between the time when an individual becomes infected and the time when that individual infects another individual \citep{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical};
therefore, estimates of \Ro from different models directly depend on their implicit assumptions about the generation-interval distribution and the exponential growth rate.
Early in an epidemic, information is scarce and there is inevitably a great deal of uncertainty surrounding both case reports (affecting the estimates of the exponential growth rate) and contact tracing (affecting the estimates of the generation-interval distribution).
We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty in \Ro should propagate error from all three sources \citep{elderd2006uncertainty}.

We compare seven disparate models published online between January 23--26, 2020 that estimated \Ro for the SARS-CoV-2 outbreak \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}. 
We use a Bayesian multilevel model to construct pooled estimates for the three key quantities: $r$, $\bar G$, and $\kappa$;
the pooled estimates reflect the uncertainties present in modeling approaches and their underlying assumptions.
We use these pooled estimates to illustrate the importance of propagating different sources of error, particularly uncertainty in both the growth rate and the generation interval. 
We also use our framework to tease apart which assumptions of these different models led to their different estimates and confidence intervals.
Despite the availability of more recent and/or updated estimates of \Ro, we restrict ourselves to the estimates above in order to focus on the resolution of uncertainty in the earliest stages of an epidemic.

@@

\begin{table}[t]
\begin{center}
\scriptsize
\begin{tabular}{l|p{3.5cm}|p{2.5cm}|p{2.7cm}|l}
 & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ & \\
\hline
Study 1 & 1.5--3.5 & 10 & 1 & \cite{bedfordncov} \\
\hline
Study 2 & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ & \cite{imaincov} \\
\hline
Study 3 & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 & \cite{liuncov} \\
\hline
Study 4 & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & \cite{readncov} \\
\hline
Study 5 & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5 & \cite{riouncov} \\
\hline
Study 6 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2 & \cite{zhaoncov} \\
\hline
Study 7 & 2.0--3.1 & 6--10 & 0 & \cite{majumderncov} \\
\hline
\end{tabular}
\end{center}
\caption{
\textbf{Reported estimates of the basic reproductive number and the assumptions about the generation-interval distributions.}
Estimates of \Ro and their assumptions about the shape of the generation interval distributions were collected from 7 studies.
$^\ast$We treat these intervals as a 95\% confidence interval in our analysis.
$^\dagger$We assume $\kappa = 0.5$ in our analysis.
$^\ddagger$The authors presented \Ro estimates under different assumptions regarding the reporting rate; we use their baseline scenario in our analysis to remain consistent with other studies, which do not account for changes in the reporting rate.
}
\end{table}

We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 7 articles that were published online between January 23--26, 2020 (Table 1).
Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the web site of Imperial College London \citep{imaincov}; and one report was posted on nextstrain.org \citep{bedfordncov}.
Their modeling approaches vary widely: a branching process model \citep{bedfordncov, imaincov, riouncov}, a deterministic Susceptible-Exposed-Infected-Recovered (SEIR) model \citep{readncov}, an exponential growth model \citep{zhaoncov}, a Poisson offspring distribution model \citep{liuncov}, and the Incidence Decay and Exponential Adjustment (IDEA) model \citep{majumderncov}.
Four studies estimated \Ro by directly fitting their models to incidence data \citep{readncov, zhaoncov, liuncov, majumderncov}.
The remaining three studies estimated \Ro by comparing the predicted number of cases from their models with the estimated number of total cases by January 18 (between 1,000 and 9,7000 \citep{imaincov0})
Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.

@@

Early in an outbreak, \Ro is difficult to estimate directly;
instead, \Ro is often inferred from the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{ma2014estimating}.
Given an estimate of the exponential growth rate $r$ and an \emph{intrinsic} generation-interval distribution $g(\tau)$ \citep{champredon2015intrinsic}, the basic reproductive
number can be estimated via the Euler-Lotka equation \citep{wallinga2007generation}:
\begin{linenomath*}
\begin{equation}
1/\Ro = \int \exp(-r\tau) g(\tau) \, d\tau.
\label{eq:euler}
\end{equation}
\end{linenomath*}
In other words, estimates of \Ro must
depend on the assumptions about the
exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.

Here, we use the gamma approximation framework \citep{park2019practical} to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
Assuming that generation intervals follow a gamma distribution with the mean $\bar G$ and the squared coefficient of variation $\kappa$, we have
\begin{linenomath*}
\begin{equation}
\Ro = \left(1 + \kappa r \bar{G}\right)^{1/\kappa}.
\label{eq:gamma}
\end{equation}
\end{linenomath*}
This equation demonstrates that a generation-interval distribution that has a larger mean (higher $\bar{G}$) or is less variable (lower $\kappa$) will give a higher estimate of \Ro for the same value of $r$ \citep{wallinga2007generation}.

@@

\newcommand{\gammdist}{\mathrm{Gamma}}
\begin{table}[t]
\begin{center}
\scriptsize
\begin{tabular}{l|p{4.5cm}|p{2.5cm}|p{2.7cm}}
 & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ \\
\hline
Study 1 & $\mathrm{Uniform}(1.5, 3.5)$ & 10 & 1 \\
\hline
Study 2 & $\gammdist(\mathrm{mean}=2.6, \alpha=28)$ & 8.4 & 0.5 \\
\hline
Study 3 & $\gammdist(\mathrm{mean}=2.92, \alpha=67)$ & 8.4 & 0.2 \\
\hline
Study 4 & $\gammdist(\mathrm{mean}=3.8, \alpha=1400)$ & 7.6 & 0.5 \\
\hline
Study 5 & $\gammdist(\mathrm{mean}=2.2, \alpha=12)$ & $\mathrm{Uniform}(7, 14)$ & 0.5\\
\hline
Study 6 & $\gammdist(\mathrm{mean}=5.47, \alpha=54)$ & $\mathrm{Uniform}(7.6, 8.4)^\ast$ & 0.2\\
\hline
Study 7 & $\exp(r \bar G)^\dagger$ & $\mathrm{Uniform}(6, 10)$ & 0\\
\hline
\end{tabular}
\end{center}
\caption{
\textbf{Probability distributions for \Ro, $\bar G$, and $\kappa$.}
We use these probability distributions to obtain a probability distribution for the exponential growth rate $r$.
The gamma distribution is parameterized by its mean and shape $\alpha$.
Constant values are fixed according to Table 1.
$^\ast$We do not account for this uncertainty during our recalculation of the exponential growth rate $r$ because the reported estimate of $\mathcal R_0$ and its uncertainty assumes $\bar G = 8$; the original article reports three $\mathcal R_0$ (and 95\% CIs) estimates using three different values of $\bar G$: 7.6 (MERS-like), 8 (average), and 8.4 (SARS-like).
We still account for this uncertainty in our pooled estimates ($\mu_G$).
$^\dagger$Study 6 uses the IDEA model \citep{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
Instead of modeling \Ro with a probability distribution and recalculating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in the reported \Ro, when combined with the considered range of $\bar G$.
}
\end{table}

@@

We use weakly informative priors on hyperparameters:
\begin{linenomath*}
\begin{equation}
\begin{aligned}
\mu_r &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week}^{-1},\,\mathrm{shape}=2)\\
\mu_G &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week},\,\mathrm{shape}=2)\\
\mu_\kappa &\sim \gammdist(\mathrm{mean}=0.5,\,\mathrm{shape}=2)\\
(\sigma_r, \sigma_G, \sigma_\kappa) &\sim \textrm{half-normal}(0, 10).
\end{aligned}
\end{equation}
\end{linenomath*}
We followed recommendations outlined in \cite{gelman2006prior}, parameterizing the top-level gamma distributions in terms of their means and standard deviations and imposing weakly informative prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
We had initially used gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) but found this put too much prior probability on large between-study variances. This phenomenon is a known problem \citep{gelman2006prior}.
Alternative choices of prior for the between-study shape parameters are also suboptimal: imposing strong priors (e.g. half-$t(\mu=0,\sigma=1,\nu=4)$  assumes \textit{a priori} that between-study variance is large,  while weak priors (e.g. half-Cauchy(0,5)) can lead to poor mixing.

We run 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 burnin steps and 500,000 sampling steps.
Posterior samples are thinned every 1000 steps.
Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \citep{gelman1992inference};
trace plots and marginal posterior distribution plots are presented in Appendix.
95\% confidence intervals are calculated by taking 2.5\% and 97.5\% quantiles from the marginal posterior distribution for each parameter.

@@

\fref{assumption} compares the reported values of the exponential growth rate $r$, mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ from different studies with the pooled estimates that we calculate from our multilevel model.
We find that there is a large uncertainty associated with the underlying parameters;
many models rely on stronger assumptions that ignore these uncertainties.
Surprisingly, no studies take into account how the variation in generation intervals affects their estimates of \Ro:
all studies assumed fixed values for $\kappa$, ranging from 0 to 1.
Assuming fixed parameter values can lead to overly strong conclusions \citep{elderd2006uncertainty}.

@@

Estimating the basic reproductive number \Ro is crucial for predicting the course of an outbreak and planning intervention strategies.
Here, we use a gamma approximation \citep{park2019practical} to decompose \Ro estimates into three key quantities ($r$, $\bar G$, and $\kappa$) and apply a multilevel Bayesian framework to compare estimates of \Ro for the novel coronavirus outbreak.
Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, including uncertainties in the amount of dispersion in the generation intervals.
Our analysis of individual studies shows that many early estimates of \Ro rely on strong assumptions.

Of the seven studies that we reviewed, two of them directly fit their models to cumulative number of confirmed cases.
This approach can be appealing because of its simplicity and apparent robustness, but fitting a model to cumulative incidence instead of raw incidence can both bias parameters and give overly narrow confidence intervals, if the resulting non-independent error structures are not taken into account \citep{ma2014estimating, king2015avoidable}.
Naive fits to cumulative incidence data should therefore be avoided.

Many sources of noise affect real-world incidence data, including both dynamical, or ``process'', noise (randomness that directly or indirectly affects disease transmission); and observation noise (randomness underlying how many of the true cases are reported).  
Disease modelers face the choice of incorporating one or both of these in their data-fitting and modeling steps. 
This is not always a serious problem, particularly if the goal is inferring parameters rather than directly making forecasts \citep{ma2014estimating}.
Modelers should however be aware of the possibility that ignoring one kind of error can give overly narrow confidence intervals \citep{king2015avoidable,taylor2016stochasticity}.

There are other important phenomena not covered by our simple framework. 
Examples that seem relevant to this outbreak include: changing reporting rates, reporting delays (including the effects of weekends and holidays), and changing generation intervals.
For emerging pathogens such as SARS-CoV-2, there may be an early period of time when the reporting rate is very low due to limited awareness or diagnostic resources;
for example, \cite{zhaoncov} (Study 6) demonstrated that estimates of \Ro can change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 17, when the official diagnostic guidelines were released \citep{who17protocol}, and January 20.
Delays between key epidemiological timings (e.g., infection, symptom onset, and detection) can also shift the shape of an observed epidemic curve and, therefore, affect parameter estimates as well as predictions of the course of an outbreak \citep{tariq2019assessing}.
Even though a constant delay between infection and detection may not affect the estimate of the growth rate, it can still affect the associated confidence intervals.
Finally, generation intervals can become shorter throughout an epidemic as intervention strategies, such as quarantine, can reduce the infectious period \citep{hethcote2002effects}.
Accounting for these factors is crucial for making accurate inferences.

Here, we focused on the estimates of \Ro that were published within a very short time frame (January 23--26).
During early phases of an outbreak, it is reasonable to assume that the epidemic grows exponentially \citep{anderson1991infectious}. However, as the number of susceptible individuals decreases, the epidemic will saturate, and estimates of $r$ used for \Ro should account for the possibility that $r$ is decreasing through time.
Although our analysis only reflects a snapshot of a fast-moving epidemic, we expect certain lessons to hold: confidence intervals must combine different sources of uncertainty. 
In fact, as epidemics progress and more data becomes available, it is likely that inferences about exponential growth rate (and other epidemiological parameters) will become more precise; thus the risk of over-confidence when uncertainty about the generation-interval distribution is neglected will become greater.

We strongly emphasize the value of attention to accurate characterization of the transmission chains via contact tracing and better statistical frameworks for inferring generation-interval distributions from such data \citep{britton2019estimation}.
A combined effort between public-health workers and modelers in this direction will be crucial for predicting the course of an epidemic and controlling it.
We also emphasize the value of transparency from modelers.
Model estimates during an outbreak, even in pre-prints, should include code links and complete explanations.
We suggest using methods based on open-source tools allow for maximal reproducibility.
